
% spMain.hlp:refs 041415

\begin{thebibliography}{99}
\bibitem{vlad}A. Vladimirescu and S. Liu, {\em The Simulation of MOS
 Integrated Circuits Using SPICE2},  ERL Memo No. ERL M80/7, Electronics
 Research Laboratory, University of  California,  Berkeley, Oct. 1980.
\bibitem{sheu}B. J. Sheu, D. L. Scharfetter, and P. K. Ko,
 {\em SPICE2 Implementation of BSIM}, ERL Memo No. ERL M85/42, Electronics
 Research Laboratory, University of California, Berkeley, May 1985.
\bibitem{pierret}J. R. Pierret, {\em A MOS Parameter Extraction Program
 for the BSIM Model},  ERL Memo Nos. ERL M84/99 and M84/100, Electronics
 Research Laboratory, University of  California, Berkeley, Nov. 1984.
\bibitem{jeng} Min-Chie Jeng, {\em Design and Modeling of Deep-Submicrometer
 MOSFETs}, ERL Memo Nos. ERL M84/99 and ERL M90/90, Electronics
 Research Laboratory, University of California, Berkeley, October 1990.
\bibitem{park}Soyeon Park, {\em Analysis and SPICE implementation of High
 Temperature Effects on MOSFET}, Master's Thesis, University of California,
 Berkeley, December 1986.
\bibitem{szeto}Clement Szeto, {\em Simulator of Temperature effects in
 MOSFETs (STEIM)}, Master's Thesis, University of California, Berkeley,
 May 1988.
\bibitem{parker}A. E. Parker and D. J. Skellern, {\em An Improved FET Model
 for Computer Simulators}, IEEE Trans. CAD, vol. 9, no.5, pp. 551-553,
 May 1990.
\bibitem{bsim3}Y. Cheng, M. Chan, K. Hui, M-C Jeng, Z. Liu, J. Huang,
 K. Chen, J. Chen, R. Tu, P. Ko and C. Hu, {\em BSIM3v3 Manual},
 Department of Electrical Engineering and Computer Sciences,
 University of California, Berkeley, 1996.
\bibitem{saleh}R. Saleh and A. Yang, Editors, {\em Simulation and Modeling,
 IEEE Circuits and Devices}, vol. 8, no. 3, pp. 7-8 and 49, May 1992. 
\bibitem{statz}H. Statz et al., {\em GaAs FET Device and Circuit
 Simulation in SPICE}, IEEE Transactions on Electron Devices, Vol. 34,
 Number 2, February 1987 pp. 160-169.
\bibitem{jewett}R. E. Jewett, {\em Josephson Junctions in SPICE2G5},
 ERL Memo, Electronics Research Laboratory, University of California,
 Berkeley, 1982.
\bibitem{whiteley}S. R. Whiteley, {\em Josephson Junctions in SPICE3},
 IEEE Trans. Magn., vol. 27, no. 2, pp. 2902-2905, March 1991.
\bibitem{ltra}J. S. Roychowdhury and D. O. Pederson, {\em Efficient
 Transient Simulation of Lossy Interconnect}, Proc. DAC 91.
\bibitem{lin}Shen Lin and Ernest S. Kuh, {\em Transient Simulation of Lossy
 Interconnect}, Proc. DAC, pp 81-86, 1992.
\bibitem{jeffery}M. Jeffery, P. Y. Xie, S. R. Whiteley, and T. Van Duzer,
  {\em Monte Carlo and thermal noise analysis of ultra-high-speed high
  temperature superconductor digital circuits},
  IEEE Trans. Applied Superconductivity,
  vol. 9, no. 2, pt. 3, pp. 4095-4098, June 1999.
\end{thebibliography}

